### 2020-05-23 chapters/Benchmark_problems.tex
iffTeX export

@Article{Pippard1957,
  Title                    = {An experimental determination of the Fermi surface in copper},
  Author                   = {Pippard, Alfred Brian},
  Journal                  = {Philosophical Transactions of the Royal Society of London. Series A, Mathematical and Physical Sciences},
  Year                     = {1957},
  Number                   = {979},
  Pages                    = {325--357},
  Volume                   = {250},

  Publisher                = {The Royal Society},
  Timestamp                = {2014.09.14},
  Url                      = {http://rsta.royalsocietypublishing.org/content/250/979/325.short}
}