### 2020-05-23 chapters/Theory.tex

\chapter{Theory}
\label{chapter:theory}
Motivated in the Chapter~\ref{chapter:intro} the starting point to study fluid flows are the Navier-Stokes equations~\cite{Navier, Stokes}.
Claude Navier in 1823 as well as Gorge G. Stokes in 1842 derived the governing equations for a fluid flow.

\section{Thin liquid films}
\label{sec:thin_films}
\begin{equation}\label{eq:simple_thin_film}
    \partial_t h = -\partial_x\left(Q(h)\gamma\partial_x^2 p\right),
\end{equation}

\section{From Shallow Water to Thin Liquid Films}
\label{sec:theory_thin_film}