### 2020-05-23 main.tex

\documentclass[titlepage,a4paper,twoside,12pt,bibliography=totoc,BCOR=6.5mm]{scrreprt}

\usepackage{main}
% use biber for build
\addbibresource{bibliography/references.bib}%

\author{Stefan Zitz}%
\title{Dynamics of thin liquid film flows studied with a lattice Boltzmann method}%
\date{\today}%

%\includeonly{chapters/example}%

\begin{document}%
%
\pagenumbering{alph}%
\pagestyle{empty}%
\include{others/titlepage}%
%
\cleardoublepage%
%
\cleardoublepage%
%
\include{others/abstract}%
%
\cleardoublepage%
%
\pagestyle{scrheadings}%
\pagenumbering{roman}%
\setcounter{page}{1}%
%
\tableofcontents%
%
\cleardoublepage%
%
\listoffigures%
%
\cleardoublepage%
%
\pagenumbering{arabic}%
\setcounter{page}{1}%
%
%
\include{chapters/Introduction}%
\include{chapters/Theory}
\include{chapters/Method}
%
%
\cleardoublepage%
\nocite{*}%
\printbibliography%
%
%
\cleardoublepage%
%
\appendix%
\renewcommand{\thepage}{\Alph{chapter}\ \arabic{page}}%
%
\setcounter{page}{1}%
\include{appendix/example}%
%
\end{document}