### 2021-06-30 chapters/Benchmark_problems.tex

\chapter{Thin Film Simulations}\label{chapter:test_problems}
\epigraph{\textit{One more time}}{Daft Punk}

\section{Rayleigh-Taylor instability}\label{sec:rayleigh-taylor}

\section{Tanner's \& Cox-Voinov law}\label{sec:droplet_relaxing}

\section{Moving droplets}\label{sec:moving drops}