### 2022-05-16 main.tex

\documentclass[titlepage,a4paper,twoside,12pt,bibliography=totoc,BCOR=6.5mm]{scrreprt}

\usepackage{main}
% use biber for build
\addbibresource{bibliography/references.bib}%

\author{Stefan Zitz}%
\title{Dynamics of thin liquid film flows studied with the lattice Boltzmann method}%
\date{\today}%

%\includeonly{chapters/example}%

\begin{document}%
%
\pagenumbering{alph}%
\pagestyle{empty}%
\include{others/titlepage}%
%
\cleardoublepage%
%
\cleardoublepage%
%
\include{others/abstract}%
%
\cleardoublepage%
%
\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\definecolor{jlblue}{rgb}{0.0,0.6056031611752245,0.9786801175696073}
\definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}

\pagestyle{scrheadings}%
\pagenumbering{roman}%
\setcounter{page}{1}%
%
\tableofcontents%
%
\cleardoublepage%
%
\listoffigures%
%
\cleardoublepage%
%
\pagenumbering{arabic}%
\setcounter{page}{1}%
%
%
\include{chapters/Introduction}%
\include{chapters/Theory}
\include{chapters/Method}
\include{chapters/Paper_four}
\include{chapters/Paper_one}
\include{chapters/Paper_two}
\include{chapters/Paper_three}
\include{chapters/Conclusion}
%
%
%
\cleardoublepage%
\nocite{*}%
\printbibliography%
%
%
\cleardoublepage
% 
\appendix%
\renewcommand{\thepage}{\Alph{chapter}\ \arabic{page}}%
%
\cleardoublepage%
%
% \setcounter{page}{1}%
\include{others/acknowledgements}
%

%\include{appendix/example}%
%
\end{document}