### 2024-01-16 appendix/List_of_publications.tex
Updates from Overleaf

\chapter{List of publications}
\label{chapter:lop}

\begin{enumerate}
    \item Zitz S, Scagliarini A, Harting J., ``Controlling the dewetting morphologies of thin liquid films by switchable substrates.'', \textit{Physical Review Fluids}, \textbf{2023} Dec 19;8(12):L122001.
    \item Zitz S, Zellhöfer M, Scagliarini A, Harting J., ``Swalbe. jl: A lattice Boltzmann solver for thin film hydrodynamics.'', \textit{The journal of open source software},  \textbf{2022} Sep 5;7(77):4312.
    \item Zitz S, Scagliarini A, Harting J., ``Lattice Boltzmann simulations of stochastic thin film dewetting.'' \textit{Physical Review E}, \textbf{2021} Sep 3;104(3):034801.
    \item Zitz S, Scagliarini A, Maddu S, Darhuber A. A., Harting J., ``Lattice Boltzmann method for thin-liquid-film hydrodynamics.'' \textit{Physical Review E}, \textbf{2019} Sep 23;100(3):033313.
\end{enumerate}