### 2024-01-16 others/secondpage.tex

\begin{tikzpicture}[overlay, remember picture]

\node at ($ (current page.north)!0.2!(current page.center) $) {%
	\parbox{\textwidth}{%
		\large{Als Dissertation genehmigt \\ von der Technischen Fakult\"at\\ der Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg}
	}%
};%

\node at ($ (current page.north)!0.32!(current page.center) $) {
    \parbox{\textwidth}{%
		\large{Tag der m\"undlichen Pr\"ufung:} 
	}%
};

\node at ($ (current page.north)!0.45!(current page.center) $) {
    \parbox{\textwidth}{%
		\large{Gutachter/in: \qquad Prof. Dr. \\ \textcolor{white}{Gutachter/in:} \qquad Prof. Dr.} 
	}%
};

\end{tikzpicture}