### 2024-01-30 chapters/chapter_1_heading.tex
Printed and evaluated hopefully

\chapter{Publications}
\label{chapter:fourth_paper}
\epigraph{\textit{Du wirst nicht enttäuscht, wenn du nie etwas erwartest und bevor du etwas falsch machst, dann mach mal lieber gar nichts.}}{Kraftklub}

The effort and work during my time as doctoral candidate has lead to four peer-reviewed articles.
In this chapter we add these articles as they appeared in their respective journal.

The first article is published in the Journal of Open Source software and discusses the code and partly the method.
The second article appeared in Physical Review E and introduces the method and the derivation of our thin film model.
The third article appeared as well in Physical Review E and discusses an extension to the thin film equation, namely a stochastic term that accounts for thermal fluctuations.
The fourth articles appeared as a letter in Physical Review Fluids and introduces the idea of a modeled switchable substrate coupled to a dewetting thin film.