### 2024-08-30 main.tex

\documentclass[titlepage,a4paper,twoside,12pt,headheight=30pt,bibliography=totoc,BCOR=6.5mm]{scrreprt}

\usepackage{main}
% use biber for build
\addbibresource{bibliography/Thesis_Swalbe.bib}%

\author{Stefan Zitz}%
\title{Lattice Boltzmann simulations of the dynamics of thin liquid films \\
       \large{Gitter Boltzmann Simulationen der Dynamik von d\"unnen Fl\"ussigkeitsfilmen}}%
\date{\today}%

%\includeonly{chapters/example}%

\begin{document}%
%
\pagenumbering{alph}%
\pagestyle{empty}%
\include{others/titlepage}%
\include{others/secondpage}
%
\cleardoublepage%
%
\cleardoublepage%
%
\include{others/abstract}%
%
\cleardoublepage%
%

\pagestyle{scrheadings}%
\pagenumbering{roman}%
\setcounter{page}{1}%
\include{others/prolog}%%
\tableofcontents%
%
\cleardoublepage%
%
% \listoffigures%
%
% \cleardoublepage%
%

\pagenumbering{arabic}%
\setcounter{page}{1}%
%
%
\include{chapters/Introduction}%
\include{chapters/Theory}
\include{chapters/Method}
\include{chapters/Conclusion}
\include{chapters/chapter_1_heading}
\includepdf[pages=-, addtotoc={1, section, 1, Swalbe.jl: A lattice Boltzmann solver for thin film hydrodynamics, paper_4}]{papers/Joss_paper.pdf}
%\include{chapters/chapter_2_heading}
\includepdf[pages=-, addtotoc={1, section, 1, Lattice Boltzmann method for thin-liquid-film hydrodynamics, paper_1}]{papers/paper_1.pdf}
%\include{chapters/chapter_3_heading}
\includepdf[pages=-, addtotoc={1, section, 1, Lattice Boltzmann simulations of stochastic thin film dewetting, paper_2}]{papers/paper_2.pdf}
%\include{chapters/chapter_4_heading}
\includepdf[pages=-, addtotoc={1, section, 1, Controlling the dewetting morphologies of thin liquid films by switchable substrates, paper_3}]{papers/paper_3.pdf}
\includepdf[pages=-]{papers/SuppMat.pdf}
\include{chapters/Paper_four}
\include{chapters/Paper_one}
\include{chapters/Paper_two}
\include{chapters/Paper_three}

%
\cleardoublepage%
\nocite{*}%
\printbibliography%
%
\cleardoublepage%
%
\appendix%
\renewcommand{\thepage}{\Alph{chapter}\ \arabic{page}}%
%
\cleardoublepage%
%
\setcounter{page}{1}%
%\include{appendix/List_of_publications}
\include{appendix/acknowledgements}
%\include{appendix/example}%
%
\end{document}